\vspace{-1.8cm} \Large Demystifying Ambient Seismic Noise Processing: \\Automating source analysis and processing decisions in the presence of coherent anthropogenic, natural, and sensor noise \vspace{-0.5cm}

### 2017-02-17 paper.tex
polishing up a few grammar errors and missing info in bibliography

\vspace{-1.8cm}
\subsection*{Motivation} 
\vspace{-0.2cm}
By measuring the speed of seismic waves propagating in the Earth's near surface, we can image the top tens to hundreds of meters of the subsurface, with deeper features being resolved by lower frequencies. These seismic velocity images can be interpreted to evaluate earthquake or landslide risk, to detect permafrost, to find sinkholes or tunnels, or to track near surface changes related to drilling activities. Additionally, in cases of complex near-surface conditions, accurate near-surface images can be used to improve deeper seismic images.
\par
By cross-correlating noise recorded at any two receivers in an array, we can extract signals mimicking an active seismic survey with a source at any receiver, virtual source response functions \cite{Lobkis2001}, \cite{Lin2008}, \cite{Wapenaar2010A}, \cite{Wapenaar2010B}. Thus, when active sources are too costly or logistically prohibitive, passive seismic can be a good option for near surface imaging. However, the theory is limited by the assumption of homogeneous uncorrelated sources \cite{Wapenaar2010A}. Non-ideal sources can cause artifacts in extracted velocities, but with careful processing experts can overcome these limitations \cite{Bensen2007}, \cite{Daskalakis2016}, \cite{Zhan2013}, even for anthropogenic sources occurring at higher frequencies \cite{Girard2016}, \cite{Martin2015}, \cite{Martin2016}, \cite{Nakata2011}, \cite{Nakata2015}, \cite{Zeng2017}. Acquisition is a further issue: nearly all of the past ambient noise studies on dense arrays have been temporary arrays which were labor-intensive to install and maintain \cite{Lin2008}, \cite{Nakata2011}, \cite{Nakata2015}, \cite{Zeng2017}. 
\par
Distributed acoustic sensing (DAS) probes a fiber optic cable with a laser interrogator unit (IU) to repurpose that fiber as a series of strain or strain rate sensors. For cost reasons, most experiments use straight fibers, but these fibers measure strain along the axis of the fiber, so some experimenters are now wrapping the fiber in a helix \cite{Kuvshinov2016}. DAS is being increasingly adopted in the energy industry for microseismic monitoring \cite{Webster2013} and time-lapse seismic surveys \cite{Daley2013}, \cite{Mateeva2014}, \cite{Bakku2015}, \cite{Miller2016} because it allows for a dense series of seismic sensors to communicate over long distances while they are run by a single power source, and those sensors can be permanently left in the same location.
\par
The maintenance cost of permanent node arrays motivated the use of buried DAS, and there have now been several ambient noise experiments where fiber optics were buried in shallow trenches so the fiber was directly coupled to the ground \cite{AjoFranklin2015}, \cite{Martin2015}, \cite{Martin2016}, \cite{Zeng2017}. In fact, we can sacrifice some ground-to-sensor coupling in favor of easier installation by running fibers in existing telecommunications conduits \cite{Martin2017}. By running fibers in existing conduits, or even plugging into existing dark fiber, on-demand repeatable seismic studies in urban areas requiring little manual labor will soon be a reality. When the interrogator unit probes a fiber, the data can serve many purposes: near surface imaging with ambient noise, earthquake detection and location (and local site response) \cite{Martin2017}, tracking vehicle traffic, potentially even detecting underground tunneling, studying the status of fluid pump systems nearby, or detecting intrusion into manholes. 
\par
With ambient noise data becoming increasingly easy to record, we can only extract its full value if we further automate the processing workflow.

@@

		\item New: clustering/separation of data (k-means, hierarchical agglomerative, SVM separation) \cite{Liao2005}, \cite{Hastie2009} to speed up noise field exploration, identification of potential issues

@@

\bibitem{AjoFranklin2015} J. Ajo-Franklin, N. Lindsey, S. Dou, T.M. Daley, B. Freifeld, E.R. Martin, M. Robertson, C. Ulrich and A. Wagner, "A Field Test of Distributed Acoustic Sensing for Ambient Noise Recording," \textit{SEG Technical Program Expanded Abstracts}, pp. 2620-2624, 2015.

@@

\bibitem{Bakku2015} S.K. Bakku, "Fracture Characterization from Seismic Measurements in a Borehole," Ph.D. Thesis, Massachusetts Institute of Technology, 2015.

@@

\bibitem{Bensen2007} G.D. Bensen, M.H. Ritzwoller, M.P. Barmin, A.L. Levshin, F. Lin, M.P. Moschetti, N.M. Shapiro and Y. Yang, "Processing seismic ambient noise data to obtain reliable broad-band surface wave dispersion measurements," \textit{Geophys. J. Int.}, vol. 169, no. ****, pp. 1239-1260, 2007.

@@

\bibitem{Daskalakis2016} E. Daskalakis, C.P. Evangelidis, J. Garnier, N.S. Melis, G. Papanicolaou and C. Tsoga, "Robust seismic velocity change estimation using ambient noise recordings," \textit{Geophys. J. Int.}, vol. 205, no. 3, pp. 1926-1936, 2016.

@@

\bibitem{Daley2013} T.M. Daley, B.M. Freifeld, J. Ajo-Franklin and S. Dou, "Field testing of fiber-optic distributed acoustic sensing (DAS) for subsurface seismic monitoring," \textit{The Leading Edge}, pp. 936-942, June 2013.

@@

\bibitem{Evans2014} J.R. Evans, R.M. Allen, A.I. Chung, E.S. Cochran, R. Guy, M. Hellweg and J.F. Lawrence, "Performance of Several Low-Cost Accelerometers," \textit{Seismol. Res. Lett.}, vol. 85, no. 1, pp. 148-158, 2014.

@@

\bibitem{Fisher2016} W.D. Fisher, T.K. Camp and V.V. Krzhizhanovskaya, "Anomaly detection in earth dam and levee passive seismic data using support vector machines and automatic feature selection," \textit{J. Comput. Sci.}, in press, doi: 10.1016/j.jocs.2016.11.016, 2016.

@@

\bibitem{Girard2016} A.J. Girard and J. Shragge, "Extracting body waves from ambient seismic recordings," \textit{SEG Technical Program Expanded Abstracts}, pp. 2715-2719, 2016.

@@

\bibitem{Kuvshinov2016} B.N. Kuvshinov, "Interaction of helically wound fibre-optic cables with plane seismic waves," \textit{Geophysical Prospecting}, vol. 64, no. 3, pp. 671-688, 2016.

@@

\bibitem{Liao2005} T.W. Liao, "Clustering of time series data- a survey," \textit{Pattern Recogn.}, vol. 38, no. 11, pp. 1857-1874, 2005.

@@

\bibitem{Lin2008} F. Lin, M.P. Moschetti and M.H. Ritzwoller, "Surface wave tomography of the western United States from ambient seismic noise: Rayleigh and Love wave phase velocity maps," \textit{Geophys. J. Int.}, vol. 173, no. ***, pp. 281-298, 2008.

@@

\bibitem{Lobkis2001} O.I. Lobkis and R.L. Weaver, "On the emergence of the Green's function in the correlations of a diffuse field," \textit{J. Acoust. Soc. Am.}, vol. 110, no. 6, pp. 3011-3017, 2001.

@@

\bibitem{Martin2015} E.R. Martin, J. Ajo-Franklin, S. Dou, N. Lindsey, T.M. Daley, B. Freifeld, M. Robertson, A. Wagner, C. Ulrich, "Interferometry of ambient noise from a trenched distributed acoustic sensing array," \textit{SEG Technical Program Expanded Abstracts}, pp. 2445-2450, 2015.

@@

\bibitem{Martin2016} E.R. Martin, N.J. Lindsey, S. Dou, J. Ajo-Franklin, T.M. Daley, B. Freifeld, M. Robertson, C. Ulrich, A. Wagner and K. Bjella, "Interferometry of a roadside DAS array in Fairbanks, AK," \textit{SEG Technical Program Expanded Abstracts}, pp. 2725-2629, 2016.

@@

\bibitem{Martin2017} E.R. Martin, B.L. Biondi, M. Karrenbach, S. Cole, "Continuous Subsurface Monitoring by Passice Seismic with Distributed Acoustic Sensors- The "Stanford Array" Experiment," \textit{Proceedings of the First EAGE Workshop on Practical Reservoir Monitoring}, 2017.

@@

\bibitem{Mateeva2014} A. Mateeva, J. Lopez, H. Potters, J. Mestayer, B. Cox, D. Kiyashchenko, P. Wills, S. Grandi, K. Hornman, B. Kuvshinov, W. Berlang, Z. Yang, R. Detomo, "Distributed acoustic sensing for reservoir monitoring with vertical seismic profiling," \textit{Geophysical Prospecting}, vol. 62, no. 4, pp. 679-692, 2014.

@@

\bibitem{Miller2016} D.E. Miller, T.M. Daley, B.M. Freifeld, M. Robertson, J. Cocker and M. Craven, "Simultaneous Acquisition of Distributed Acoustic Sensing VSP with Multi-mode and Single-mode FIber-optic Cables and 3C-Geophones at the Aquistore CO$_2$ Storage Site," \textit{CSEG Recorder}, pp. 28-***, June 2016.

@@

\bibitem{Nakata2011} N. Nakata, R. Snieder, T. Tsuji, K. Larner and T. Matsuoka, "Shear wave imaging from traffic noise using seismic interferometry by cross-coherence," \textit{Geophysics}, vol. 76, no. 6, pp. SA97-SA106, 2011.

@@

\bibitem{Nakata2015} N. Nakata, J.P. Chang, J.F. Lawrence and P. Bou\'{e}, "Body wave extraction and tomography at Long Beach, California, with ambient-noise interferometry," \textit{J. Geophys. Res. Solid Earth}, vol. 120, pp. 1159-1173, 2015.

@@

\bibitem{Posey2000} R. Posey Jr., G.A. Johnson and S.T. Vohra, "Strain sensing based on coherence Rayleigh scattering in an optical fibre," \textit{Electronics Letters}, vol. 36, no. 20, pp. 1688-1689, 2000.

@@

\bibitem{Seats2012} K.J. Seats, J.F. Lawrence and G.A. Prieto, "Improved ambient noise correlation functions using Welch's method," \textit{Geophys. J. Int.}, vol. 188, no. ***, pp. 513-523, 2012.

@@

\bibitem{Wapenaar2010A} K. Wapenaar, D. Draganov, R. Snieder, X. Campman and A. Verdel, "Tutorial on seismic interferometry: Part 1- Basic principles and applications," \textit{Geophysics}, vol. 75, no. 5, pp. 75A195-75A209, 2010.

@@

\bibitem{Wapenaar2010B} K. Wapenaar, E. Slob, R. Snieder and A. Curtis, "Tutorial on seismic interferometry: Part 2 - Underlying theory and new advances," \textit{Geophysics}, vol. 75, no. 5, pp. 75A211-75A227, 2010.

@@

\bibitem{Webster2013} P. Webster, J. Wall, C. Perkins and M. Molenaar, "Micro-Seismic Detection using Distributed Acoustic Sensing," \textit{SEG Technical Program Expanded Abstracts}, pp. 2459-2463, 2013.

@@

\bibitem{Withers1998} M. Withers, R. Aster, C. Young, J. Beiriger, M. Harris, S. Moore and J. Trujillo, "A comparison of select trigger algorithms for automated global seismic phase and event detection," \textit{B. Seismol. Soc. Am.}, vol. 88, no. ****, pp. 95-106, 1998.

@@

\bibitem{Yoon2015} C.E. Yoon, O. O'Reilly, K.J. Bergen and G.C. Beroza, "Earthquake detection through computationally efficient similarity search," \textit{Science Advances}, vol. 1, no. ****, pp. e1501057-e1501057, 2015.

@@

\bibitem{Zeng2017} X. Zeng, C. Thurber, H. Wang, D. Fratta, E. Matzel, PoroTomo Team, "High-resolution Shallow Structure Revealed with Ambient Noise Tomography on a Dense Array," \textit{Proceedings, 42nd Workshop on Geothermal Reservoir Engineering}, Stanford University, CA, Feb. 13-15, 2017, SGP-TR-212.

@@

\bibitem{Zhan2013} Z. Zhan, V.C. Tsai and R.W. Clayton, "Spurious velocity changes caused by temporal variations in ambient noise frequency content," \textit{Geophys. J. Int}, vol. 194, no. ***, pp. 1574-1581, 2013.

>>>>

\vspace{-1.8cm}
\subsection*{Motivation} 
\vspace{-0.2cm}
By measuring the speed of seismic waves propagating in the Earth's near surface, we can image the top tens to hundreds of meters of the subsurface, with deeper features being resolved by lower frequencies. These seismic velocity images can be interpreted to evaluate earthquake or landslide risk, to detect permafrost, to find sinkholes or tunnels, or to track near surface changes related to drilling activities. Additionally, in cases of complex near-surface conditions, accurate near-surface images can be used to improve deeper seismic images.
\par
By cross-correlating noise recorded at any two receivers in an array, we can extract signals mimicking an active seismic survey with a source at any receiver, called virtual source response functions \cite{Lobkis2001}, \cite{Lin2008}, \cite{Wapenaar2010A}, \cite{Wapenaar2010B}. Thus, when active sources are too costly or logistically prohibitive, passive seismic can be a good option for near surface imaging. However, the theory is limited by the assumption of homogeneous uncorrelated sources \cite{Wapenaar2010A}. Non-ideal sources can cause artifacts in extracted velocities, but with careful processing experts can overcome these limitations \cite{Bensen2007}, \cite{Daskalakis2016}, \cite{Zhan2013}, even for anthropogenic sources occurring at higher frequencies \cite{Girard2016}, \cite{Martin2015}, \cite{Martin2016}, \cite{Nakata2011}, \cite{Nakata2015}, \cite{Zeng2017}. Acquisition is a further issue: nearly all of the past ambient noise studies on dense arrays have been temporary arrays which were labor-intensive to install and maintain \cite{Lin2008}, \cite{Nakata2011}, \cite{Nakata2015}, \cite{Zeng2017}. 
\par
Distributed acoustic sensing (DAS) probes a fiber optic cable with a laser interrogator unit (IU) to repurpose that fiber as a series of strain or strain rate sensors. For cost reasons, most experiments use straight fibers, but these fibers measure strain along the axis of the fiber, so some experimenters are now wrapping the fiber in a helix \cite{Kuvshinov2016}. DAS is being increasingly adopted in the energy industry for microseismic monitoring \cite{Webster2013} and time-lapse seismic surveys \cite{Daley2013}, \cite{Mateeva2014}, \cite{Bakku2015}, \cite{Miller2016} because it allows for a dense series of seismic sensors to communicate over long distances while they are run by a single power source, and those sensors can be permanently left in the same location.
\par
The maintenance cost of permanent node arrays motivated the use of buried DAS, and there have now been several ambient noise experiments where fiber optics were buried in shallow trenches so the fiber was directly coupled to the ground \cite{AjoFranklin2015}, \cite{Martin2015}, \cite{Martin2016}, \cite{Zeng2017}. In fact, we can sacrifice some ground-to-sensor coupling in favor of easier installation by running fibers in existing telecommunications conduits \cite{Martin2017}. By running fibers in existing conduits, or even plugging into existing dark fiber, on-demand repeatable seismic studies in urban areas requiring little manual labor will soon be a reality. When the interrogator unit probes a fiber, the data can serve many purposes: near surface imaging with ambient noise, earthquake detection and location (and local site response) \cite{Martin2017}, tracking vehicle traffic, potentially even detecting underground tunneling, studying the status of fluid pump systems nearby, or detecting intrusion into manholes. 
\par
With ambient noise data becoming increasingly easy to record, we can only extract its full value if we further automate the processing workflow.

@@

		\item New: clustering/separation of data with features including wavelet transforms (k-means, hierarchical agglomerative, SVM separation) \cite{Liao2005}, \cite{Hastie2009}, \cite{Mallat2008} to speed up noise field exploration, identification of potential issues

@@

\bibitem{AjoFranklin2015} J. Ajo-Franklin, N. Lindsey, S. Dou, T.M. Daley, B. Freifeld, E.R. Martin, M. Robertson, C. Ulrich and A. Wagner, ``A Field Test of Distributed Acoustic Sensing for Ambient Noise Recording," \textit{SEG Technical Program Expanded Abstracts}, pp. 2620-2624, 2015.

@@

\bibitem{Bakku2015} S.K. Bakku, ``Fracture Characterization from Seismic Measurements in a Borehole," Ph.D. Thesis, Massachusetts Institute of Technology, 2015.

@@

\bibitem{Bensen2007} G.D. Bensen, M.H. Ritzwoller, M.P. Barmin, A.L. Levshin, F. Lin, M.P. Moschetti, N.M. Shapiro and Y. Yang, ``Processing seismic ambient noise data to obtain reliable broad-band surface wave dispersion measurements," \textit{Geophys. J. Int.}, vol. 169, no. 3, pp. 1239-1260, 2007.

@@

\bibitem{Daskalakis2016} E. Daskalakis, C.P. Evangelidis, J. Garnier, N.S. Melis, G. Papanicolaou and C. Tsoga, ``Robust seismic velocity change estimation using ambient noise recordings," \textit{Geophys. J. Int.}, vol. 205, no. 3, pp. 1926-1936, 2016.

@@

\bibitem{Daley2013} T.M. Daley, B.M. Freifeld, J. Ajo-Franklin and S. Dou, ``Field testing of fiber-optic distributed acoustic sensing (DAS) for subsurface seismic monitoring," \textit{The Leading Edge}, pp. 936-942, June 2013.

@@

\bibitem{Evans2014} J.R. Evans, R.M. Allen, A.I. Chung, E.S. Cochran, R. Guy, M. Hellweg and J.F. Lawrence, ``Performance of Several Low-Cost Accelerometers," \textit{Seismol. Res. Lett.}, vol. 85, no. 1, pp. 148-158, 2014.

@@

\bibitem{Fisher2016} W.D. Fisher, T.K. Camp and V.V. Krzhizhanovskaya, ``Anomaly detection in earth dam and levee passive seismic data using support vector machines and automatic feature selection," \textit{J. Comput. Sci.}, in press, doi: 10.1016/j.jocs.2016.11.016, 2016.

@@

\bibitem{Girard2016} A.J. Girard and J. Shragge, ``Extracting body waves from ambient seismic recordings," \textit{SEG Technical Program Expanded Abstracts}, pp. 2715-2719, 2016.

@@

\bibitem{Kuvshinov2016} B.N. Kuvshinov, ``Interaction of helically wound fibre-optic cables with plane seismic waves," \textit{Geophysical Prospecting}, vol. 64, no. 3, pp. 671-688, 2016.

@@

\bibitem{Liao2005} T.W. Liao, ``Clustering of time series data- a survey," \textit{Pattern Recogn.}, vol. 38, no. 11, pp. 1857-1874, 2005.

@@

\bibitem{Lin2008} F. Lin, M.P. Moschetti and M.H. Ritzwoller, ``Surface wave tomography of the western United States from ambient seismic noise: Rayleigh and Love wave phase velocity maps," \textit{Geophys. J. Int.}, vol. 173, no. 1, pp. 281-298, 2008.

@@

\bibitem{Lobkis2001} O.I. Lobkis and R.L. Weaver, ``On the emergence of the Green's function in the correlations of a diffuse field," \textit{J. Acoust. Soc. Am.}, vol. 110, no. 6, pp. 3011-3017, 2001.

@@

\bibitem{Martin2015} E.R. Martin, J. Ajo-Franklin, S. Dou, N. Lindsey, T.M. Daley, B. Freifeld, M. Robertson, A. Wagner, C. Ulrich, ``Interferometry of ambient noise from a trenched distributed acoustic sensing array," \textit{SEG Technical Program Expanded Abstracts}, pp. 2445-2450, 2015.

@@

\bibitem{Martin2016} E.R. Martin, N.J. Lindsey, S. Dou, J. Ajo-Franklin, T.M. Daley, B. Freifeld, M. Robertson, C. Ulrich, A. Wagner and K. Bjella, ``Interferometry of a roadside DAS array in Fairbanks, AK," \textit{SEG Technical Program Expanded Abstracts}, pp. 2725-2629, 2016.

@@

\bibitem{Martin2017} E.R. Martin, B.L. Biondi, M. Karrenbach, S. Cole, ``Continuous Subsurface Monitoring by Passice Seismic with Distributed Acoustic Sensors- The `Stanford Array' Experiment," \textit{Proceedings of the First EAGE Workshop on Practical Reservoir Monitoring}, 2017.

@@

\bibitem{Mateeva2014} A. Mateeva, J. Lopez, H. Potters, J. Mestayer, B. Cox, D. Kiyashchenko, P. Wills, S. Grandi, K. Hornman, B. Kuvshinov, W. Berlang, Z. Yang, R. Detomo, ``Distributed acoustic sensing for reservoir monitoring with vertical seismic profiling," \textit{Geophysical Prospecting}, vol. 62, no. 4, pp. 679-692, 2014.

@@

\bibitem{Miller2016} D.E. Miller, T.M. Daley, B.M. Freifeld, M. Robertson, J. Cocker and M. Craven, ``Simultaneous Acquisition of Distributed Acoustic Sensing VSP with Multi-mode and Single-mode FIber-optic Cables and 3C-Geophones at the Aquistore CO$_2$ Storage Site," \textit{CSEG Recorder}, vol. 41, no. 6, pp. 28-33, 2016.

@@

\bibitem{Nakata2011} N. Nakata, R. Snieder, T. Tsuji, K. Larner and T. Matsuoka, ``Shear wave imaging from traffic noise using seismic interferometry by cross-coherence," \textit{Geophysics}, vol. 76, no. 6, pp. SA97-SA106, 2011.

@@

\bibitem{Nakata2015} N. Nakata, J.P. Chang, J.F. Lawrence and P. Bou\'{e}, ``Body wave extraction and tomography at Long Beach, California, with ambient-noise interferometry," \textit{J. Geophys. Res. Solid Earth}, vol. 120, pp. 1159-1173, 2015.

@@

\bibitem{Posey2000} R. Posey Jr., G.A. Johnson and S.T. Vohra, ``Strain sensing based on coherence Rayleigh scattering in an optical fibre," \textit{Electronics Letters}, vol. 36, no. 20, pp. 1688-1689, 2000.

@@

\bibitem{Seats2012} K.J. Seats, J.F. Lawrence and G.A. Prieto, ``Improved ambient noise correlation functions using Welch's method," \textit{Geophys. J. Int.}, vol. 188, no. 2, pp. 513-523, 2012.

@@

\bibitem{Wapenaar2010A} K. Wapenaar, D. Draganov, R. Snieder, X. Campman and A. Verdel, ``Tutorial on seismic interferometry: Part 1- Basic principles and applications," \textit{Geophysics}, vol. 75, no. 5, pp. 75A195-75A209, 2010.

@@

\bibitem{Wapenaar2010B} K. Wapenaar, E. Slob, R. Snieder and A. Curtis, ``Tutorial on seismic interferometry: Part 2 - Underlying theory and new advances," \textit{Geophysics}, vol. 75, no. 5, pp. 75A211-75A227, 2010.

@@

\bibitem{Webster2013} P. Webster, J. Wall, C. Perkins and M. Molenaar, ``Micro-Seismic Detection using Distributed Acoustic Sensing," \textit{SEG Technical Program Expanded Abstracts}, pp. 2459-2463, 2013.

@@

\bibitem{Withers1998} M. Withers, R. Aster, C. Young, J. Beiriger, M. Harris, S. Moore and J. Trujillo, ``A comparison of select trigger algorithms for automated global seismic phase and event detection," \textit{B. Seismol. Soc. Am.}, vol. 88, no. 1, pp. 95-106, 1998.

@@

\bibitem{Yoon2015} C.E. Yoon, O. O'Reilly, K.J. Bergen and G.C. Beroza, ``Earthquake detection through computationally efficient similarity search," \textit{Science Advances}, vol. 1, no. 11, pp. e1501057, 2015.

@@

\bibitem{Zeng2017} X. Zeng, C. Thurber, H. Wang, D. Fratta, E. Matzel, PoroTomo Team, ``High-resolution Shallow Structure Revealed with Ambient Noise Tomography on a Dense Array," \textit{Proceedings, 42nd Workshop on Geothermal Reservoir Engineering}, Stanford University, CA, Feb. 13-15, 2017, SGP-TR-212.

@@

\bibitem{Zhan2013} Z. Zhan, V.C. Tsai and R.W. Clayton, ``Spurious velocity changes caused by temporal variations in ambient noise frequency content," \textit{Geophys. J. Int}, vol. 194, no. 3, pp. 1574-1581, 2013.
